---
title: "Stats desc 2015"
author: "Noémie Guibe"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```


```{r}
pop2015 <- read_csv(here::here("data/pop2015.csv"))
head(pop2015)
```

<br>

```{r}
table(pop2015$rf_artec)

round(table(pop2015$rf_fedrg)/length(pop2015$rf_fedrg),3) # soit 45% 

table(pop2015$rf_artec)

round(table(pop2015$rf_artec)/length(pop2015$rf_artec),3) # soit 55%

# recoupement absent?

round(table(pop2015$priorlive)/length(pop2015$priorlive),3)
```

<br>

Parmi ceux qui ont eu recours à des traitements contre l'infertilité:

- 27722 ont recours à des traitements contre l'infertilité sous forme de médicaments ou insémination artificielle;
- 34235 sous forme d'assisted reproductive technology (ART) (IVF ou autres)

De plus, 56% n'ont pas eu d'enfant auparavant

<br>


## AGE DES PARENTS

Age de la mère : 
12 10 – 12 years
13 13 years
...
49 49 years
50 50 years and over

Age du père :
01 Under 15 years
02 15-19 years
03 20-24 years
04 25-29 years
05 30-34 years
06 35-39 years
07 40-44 years
08 45-49 years
09 50-54 years
10 55-98 years
11 Not stated

```{r}
summary(select(pop2015,mager))
summary(select(pop2015,fagerec11))
round(table(pop2015$fagerec11)/length(pop2015$fagerec11),3)
```

<br>

# 60% des pères sont âgés de 30 à 39 ans

<br>

```{r}
plot(density(pop2015$mager), lwd = 2, col = "orange", xlab = "", main = "âge de la mère à la naissance")
hist(pop2015$fagerec11, col = "cornflowerblue", xlab = "", main  = "âge du père à la naissance")
```

<br>

# ETHNIE DES PARENTS

Race de la mère : 
1 White (only)
3 AIAN (American Indian or Alaskan Native) (only)
4 Asian (only)
5 NHOPI (Native Hawaiian or Other Pacific Islander) (only)
6 More than one race

Ethnie du père : 1.34, 75% blanc
1 White (only)
2 Black (only)
3 AIAN (only)
4 Asian (only)
5 NHOPI (only)
6 More than one race
9 Unknown or Not Stated

<br>

```{r}
summary(select(pop2015,mrace6))
summary(select(pop2015,frace6))
round(table(pop2015$frace6)/length(pop2015$frace6),3)
# 85 % des pères sont blancs, 9% asiatiques et 5% noirs
```

```{r}
hist(pop2015$frace6, col = "cornflowerblue", xlab = "", main  = "ethnie du père")
```

<br>

# EDUCATION DES PARENTS
éducation de la mère : 
1 8th grade or less
2 9th through 12th grade with no diploma
3 High school graduate or GED completed
4 Some college credit, but not a degree.
5 Associate degree (AA,AS)
6 Bachelor’s degree (BA, AB, BS)
7 Master’s degree (MA, MS, MEng, MEd, MSW, MBA)
8 Doctorate (PhD, EdD) or Professional Degree (MD, DDS,DVM, LLB, JD)
9 Unknown

Education du père : 5.6, 
1 8th grade or less
2 9th through 12th grade with no diploma
3 High school graduate or GED completed
4 Some college credit, but not a degree.
5 Associate degree (AA,AS)
6 Bachelor’s degree (BA, AB, BS)
7 Master’s degree (MA, MS, MEng, MEd, MSW, MBA)
8 Doctorate (PhD, EdD) or Professional Degree (MD, DDS,DVM, LLB, JD)
9 Unknown


```{r}
summary(select(pop2015,meduc))
summary(select(pop2015,feduc))
round(table(pop2015$meduc)/length(pop2015$meduc),3)
round(table(pop2015$feduc)/length(pop2015$feduc),3)
```


```{r}
hist(pop2015$feduc, col = "cornflowerblue", xlab = "", main  = "niveau d'éducation du père")
hist(pop2015$meduc, col = "cornflowerblue", xlab = "", main  = "niveau d'éducation de la mère")
```

<br>

# CARECTERISTIQUES DU COUPLE

Statut marital : 
1 Married
2 Unmarried

Enfant(s) encore en vie : 
00-30 Number of children still living from previous live births.
99 Unknown or not stated


```{r}
summary(select(pop2015,dmar))
summary(select(pop2015,priorlive))
round(table(pop2015$dmar)/length(pop2015$dmar),3) # 97% sont mariés
round(table(pop2015$priorlive)/length(pop2015$priorlive),3) # 87% ont déjà au plus un enfant
```

```{r}
hist(pop2015$priorlive, col = "cornflowerblue", xlab = "", main  = "nombre d'enfants")
```

<br>

# SANTE DE L'ENFANT

Score Apgar : 3.76, 50% au moins 9
1 A score of 0-3
2 A score of 4-6
3 A score of 7-8
4 A score of 9-10
5 Unknown or not stated

Poids à la naissance :
0227-8165 Number of grams
9999 Not stated birth weight


```{r}
summary(select(pop2015,apgar5r))
summary(select(pop2015,dbwt))
```

```{r}
hist(pop2015$apgar5r, col = "cornflowerblue", xlab = "", main  = "Score Apgar")
```



```{r}
plot(density(pop2015$dbwt), lwd = 2, col = "orange", xlab = "", main = "poids à la naissance")
```


